The ppmData R-package for setting up spatial point process models

### 2022-09-08 paper/paper.Rmd
beta hunting
update

```{=tex}
\begin{align}
\ell(\bm{\beta}|\bm{y}) & = \sum_{i=1}^n\log(\lambda_i) - \int_{\bm{s} \in \mathcal{A}}\lambda(\bm{s})d\bm{s} - \log(n!) \nonumber\\
& \approx \sum_{i=1}^n\log(\lambda_i) - \sum_{i=1}^m w_i\lambda_i \nonumber\\
& = \sum_{i=1}^m w_i\{z_i \log(\lambda_i) - \lambda_i\} \numberthis \label{eq:two}
\end{align}
```
where $z_i$ is an indicator variable identifying if site $i$ is one of the $n$ observed points from the random point pattern or one of the $m$ quadrature points. $w_i$ stores the quadrature weights which sum to the total area of the region $|\mathcal{A}|$. In \eqref{eq:two}, we can see the integral in the log-likelihood $\ell(\beta|\mathbf{y})$ is being estimated using quadrature [@baddeley_practical_2000; @diggle_geostatistical_2010]. @warton_poisson_2010 demonstrated for presence-only species distribution models, that using numerical quadrature with a point process framework makes the models scale invariant and treats the quadrature purely as a tool for approximating the integral.

@@

```{r fit model, cache=TRUE, warning=FALSE, message=FALSE}
#load relevant libraries
library(ppmData)
library(terra) ## need this load rasters
#read the data into R
path <- system.file("extdata", package = "ppmData")
lst <- list.files(path=path,pattern='*.tif',full.names = TRUE)
preds <- rast(lst)
#Format the data (subset and scale) 
presences <- subset(snails,SpeciesID %in% "Tasmaphena sinclairi")
preds <- scale(preds)
#Obtain quadrature points, variables and weights
ppmdata <- ppmData(npoints = 20000, presences = presences,
                   window = preds[[1]], covariates = preds)
```

>>>>

```{=tex}
\begin{align}
\ell(\bm{{\beta}}|\bm{y}) & = \sum_{i=1}^n\log(\lambda_i) - \int_{\bm{s} \in \mathcal{A}}\lambda(\bm{s})d\bm{s} - \log(n!) \nonumber\\
& \approx \sum_{i=1}^n\log(\lambda_i) - \sum_{i=1}^m w_i\lambda_i \nonumber\\
& = \sum_{i=1}^m w_i\{z_i \log(\lambda_i) - \lambda_i\} \numberthis \label{eq:two}
\end{align}
```
where $z_i$ is an indicator variable identifying if site $i$ is one of the $n$ observed points from the random point pattern or one of the $m$ quadrature points. $w_i$ stores the quadrature weights which sum to the total area of the region $|\mathcal{A}|$. In \eqref{eq:two}, we can see the integral in the log-likelihood $\ell(\beta|\mathbf{y})$ is being estimated using quadrature [@baddeley_practical_2000; @diggle_geostatistical_2010]. @warton_poisson_2010 demonstrated for presence-only species distribution models, that using numerical quadrature with a point process framework makes the models scale invariant and treats the quadrature purely as a tool for approximating the integral.

@@

```{r fit model, cache=TRUE, warning=FALSE, message=FALSE}
#load relevant libraries
library(ppmData)
#read the data into R
path <- system.file("extdata", package = "ppmData")
lst <- list.files(path=path,pattern='*.tif',full.names = TRUE)
preds <- rast(lst)
#Format the data (subset and scale) 
presences <- subset(snails,SpeciesID %in% "Tasmaphena sinclairi")
preds <- scale(preds)
#Obtain quadrature points, variables and weights
ppmdata <- ppmData(npoints = 20000, presences = presences,
                   window = preds[[1]], covariates = preds)
```
